---
title: "Scenario Comparison Notebook"
format:
  html:
    toc: true
    code-fold: true
execute:
  message: false
  warning: false
editor: visual
---

## Introduction

This notebook compares multiple simulation scenarios in one run. Edit only the **User Settings** chunk, then click **Run All** or **Render** in RStudio. Each scenario can use different true probability settings. The final output is a set of aggregate comparison tables.

Use quick mode for a fast smoke test with 5 trial simulations per scenario. Switch `quick_mode` to `FALSE` for the production default of 2,000 trial simulations per scenario.

## User Settings

Edit this chunk only for routine use.

```{r user_settings}
quick_mode <- FALSE
dose_levels <- c(1, 2, 3, 4, 5)

comparison_settings <- list(
  dose_levels = dose_levels,
  n_stages = if (quick_mode) 3 else 5,
  n_simulations = if (quick_mode) 5 else 2000,
  cohort_size = 15,
  phi_T = 0.30,
  c_T = 0.35,
  phi_E = 0.25,
  c_E = 0.60,
  phi_I = 0.20,
  c_I = 0.50,
  c_poc = 0.90,
  delta_poc = 0.8,
  use_calibration_results = TRUE,
  threshold_calibration_results_path = "results/threshold_calibration/threshold_calibration_results.rds",
  poc_calibration_results_path = "results/notebook_calibration/poc_calibration_results.rds",
  seed = 11118
)

scenarios <- list(
  list(
    name = "Increasing immune and efficacy",
    p_YI = c(0.10, 0.30, 0.50, 0.60, 0.70),
    p_YT_given_I = matrix(c(
      0.05, 0.10, 0.12, 0.18, 0.25,
      0.08, 0.12, 0.15, 0.25, 0.35
    ), nrow = length(dose_levels), ncol = 2),
    p_YE_given_I = matrix(c(
      0.10, 0.20, 0.35, 0.45, 0.50,
      0.30, 0.50, 0.70, 0.80, 0.75
    ), nrow = length(dose_levels), ncol = 2),
    rho0 = 0,
    rho1 = 0
  ),
  list(
    name = "Middle dose best",
    p_YI = c(0.08, 0.22, 0.45, 0.55, 0.60),
    p_YT_given_I = matrix(c(
      0.04, 0.08, 0.15, 0.25, 0.40,
      0.06, 0.10, 0.20, 0.32, 0.50
    ), nrow = length(dose_levels), ncol = 2),
    p_YE_given_I = matrix(c(
      0.08, 0.22, 0.40, 0.38, 0.30,
      0.25, 0.48, 0.75, 0.70, 0.55
    ), nrow = length(dose_levels), ncol = 2),
    rho0 = 0,
    rho1 = 0
  ),
  list(
    name = "Flat null",
    p_YI = rep(0.20, length(dose_levels)),
    p_YT_given_I = matrix(c(
      0.15, 0.15, 0.15, 0.15, 0.15,
      0.15, 0.15, 0.15, 0.15, 0.15
    ), nrow = length(dose_levels), ncol = 2),
    p_YE_given_I = matrix(c(
      0.20, 0.20, 0.20, 0.20, 0.20,
      0.20, 0.20, 0.20, 0.20, 0.20
    ), nrow = length(dose_levels), ncol = 2),
    rho0 = 0,
    rho1 = 0
  )
)
```

## Internal Setup

This chunk loads the backend code automatically.

```{r setup}
#| echo: false
library(knitr)

project_root_candidates <- c(".", "..", "../..")
project_root_matches <- project_root_candidates[
  file.exists(file.path(project_root_candidates, "DoseFinding.Rproj"))
]
if (length(project_root_matches) == 0) {
  stop("Could not find project root containing DoseFinding.Rproj.")
}
project_root <- normalizePath(project_root_matches[[1]], winslash = "/", mustWork = TRUE)
project_root_abs <- project_root
setwd(project_root)
cat("Project root:", project_root, "\n")

source(file.path(project_root, "src/core/config.R"))
source(file.path(project_root, "src/utils/helpers.R"))
source(file.path(project_root, "src/utils/plotting_extensions.R"))
source(file.path(project_root, "src/core/simulate_data.R"))
source(file.path(project_root, "src/core/model_utils.R"))
source(file.path(project_root, "src/decision/dose_decision.R"))
source(file.path(project_root, "src/core/main.R"))

resolve_project_path <- function(path) {
  if (is.null(path) || length(path) == 0 || is.na(path) || path == "") {
    return(path)
  }
  if (grepl("^(/|[A-Za-z]:[\\\\/])", path)) {
    return(path)
  }
  file.path(project_root_abs, path)
}

comparison_settings$threshold_calibration_results_path <- resolve_project_path(
  comparison_settings$threshold_calibration_results_path
)
comparison_settings$poc_calibration_results_path <- resolve_project_path(
  comparison_settings$poc_calibration_results_path
)
```

## Calibrated Inputs

This chunk reads saved calibration results when available.

```{r calibrated_inputs}
#| echo: false
calibration_source <- c(
  c_T = "User Settings fallback",
  c_I = "User Settings fallback",
  c_E = "User Settings fallback",
  c_poc = "User Settings fallback"
)

apply_scalar_calibration_value <- function(settings, sources, name, value, source_label) {
  if (!is.numeric(value) || length(value) != 1 || is.na(value)) {
    stop("Invalid calibration value for ", name, ".")
  }
  settings[[name]] <- as.numeric(value)
  sources[[name]] <- source_label
  list(settings = settings, sources = sources)
}

if (isTRUE(comparison_settings$use_calibration_results)) {
  threshold_path <- comparison_settings$threshold_calibration_results_path
  if (file.exists(threshold_path)) {
    threshold_results <- readRDS(threshold_path)
    recommended_thresholds <- threshold_results$recommended_thresholds
    required_thresholds <- c("c_T", "c_I", "c_E")
    missing_thresholds <- setdiff(required_thresholds, names(recommended_thresholds))
    if (length(missing_thresholds) > 0) {
      stop(
        "Threshold calibration results are missing: ",
        paste(missing_thresholds, collapse = ", ")
      )
    }
    for (name in required_thresholds) {
      applied <- apply_scalar_calibration_value(
        comparison_settings,
        calibration_source,
        name,
        recommended_thresholds[[name]],
        paste0("Threshold calibration results: ", threshold_path)
      )
      comparison_settings <- applied$settings
      calibration_source <- applied$sources
    }
  } else {
    warning(
      "Threshold calibration results were not found at ",
      threshold_path,
      ". Using c_T, c_I, and c_E from User Settings."
    )
  }

  poc_path <- comparison_settings$poc_calibration_results_path
  if (file.exists(poc_path)) {
    poc_results <- readRDS(poc_path)
    applied <- apply_scalar_calibration_value(
      comparison_settings,
      calibration_source,
      "c_poc",
      poc_results$optimal_c_poc,
      paste0("PoC calibration results: ", poc_path)
    )
    comparison_settings <- applied$settings
    calibration_source <- applied$sources
  } else {
    warning(
      "PoC calibration results were not found at ",
      poc_path,
      ". Using c_poc from User Settings."
    )
  }
}

calibration_input_table <- data.frame(
  parameter = names(calibration_source),
  current_value = vapply(names(calibration_source), function(name) {
    comparison_settings[[name]]
  }, numeric(1)),
  source = unname(calibration_source),
  stringsAsFactors = FALSE
)

kable(calibration_input_table, caption = "Calibration values used for scenario comparison")
```

## Internal Configuration

This chunk converts the user settings into the backend trial configuration.

```{r config}
#| echo: false
comparison_trial_config <- list(
  dose_levels = comparison_settings$dose_levels,
  n_stages = comparison_settings$n_stages,
  cohort_size = comparison_settings$cohort_size,
  phi_T = comparison_settings$phi_T,
  c_T = comparison_settings$c_T,
  phi_E = comparison_settings$phi_E,
  c_E = comparison_settings$c_E,
  phi_I = comparison_settings$phi_I,
  c_I = comparison_settings$c_I,
  c_poc = comparison_settings$c_poc,
  delta_poc = comparison_settings$delta_poc,
  enable_early_termination = TRUE,
  log_early_termination = FALSE,
  verbose_logging = FALSE
)

utility_table <- array(0, dim = c(2, 2, 2), dimnames = list(
  E = c(0, 1),
  T = c(0, 1),
  I = c(0, 1)
))

utility_table[1, 1, 1] <- 0
utility_table[2, 1, 1] <- 80
utility_table[1, 2, 1] <- 0
utility_table[2, 2, 1] <- 30

utility_table[1, 1, 2] <- 10
utility_table[2, 1, 2] <- 100
utility_table[1, 2, 2] <- 0
utility_table[2, 2, 2] <- 40

comparison_trial_config$utility_table <- utility_table
```

## Scenario Check

This chunk checks that every scenario has valid probability inputs.

```{r scenario_check}
#| echo: false
format_prob_vector <- function(values) {
  paste(sprintf("%.2f", values), collapse = ", ")
}

calculate_marginal <- function(prob_given_i, p_i) {
  prob_given_i[, 1] * (1 - p_i) + prob_given_i[, 2] * p_i
}

calculate_true_utilities <- function(scenario, dose_levels, utility_table) {
  vapply(seq_along(dose_levels), function(dose_index) {
    calculate_utility_from_true_probs(
      dose_idx = dose_index,
      p_YI = scenario$p_YI,
      p_YT_given_I = scenario$p_YT_given_I,
      p_YE_given_I = scenario$p_YE_given_I,
      utility_table = utility_table
    )
  }, numeric(1))
}

validate_scenario <- function(scenario, scenario_index, dose_levels) {
  required_names <- c("name", "p_YI", "p_YT_given_I", "p_YE_given_I", "rho0", "rho1")
  missing_names <- setdiff(required_names, names(scenario))
  if (length(missing_names) > 0) {
    stop("Scenario ", scenario_index, " is missing: ", paste(missing_names, collapse = ", "))
  }

  n_doses <- length(dose_levels)
  if (length(scenario$p_YI) != n_doses) {
    stop("Scenario ", scenario$name, " has p_YI length ", length(scenario$p_YI), "; expected ", n_doses, ".")
  }
  if (!all(dim(scenario$p_YT_given_I) == c(n_doses, 2))) {
    stop("Scenario ", scenario$name, " must have p_YT_given_I dimension ", n_doses, " x 2.")
  }
  if (!all(dim(scenario$p_YE_given_I) == c(n_doses, 2))) {
    stop("Scenario ", scenario$name, " must have p_YE_given_I dimension ", n_doses, " x 2.")
  }

  probability_values <- c(scenario$p_YI, scenario$p_YT_given_I, scenario$p_YE_given_I)
  if (any(is.na(probability_values)) || any(probability_values < 0) || any(probability_values > 1)) {
    stop("Scenario ", scenario$name, " has probabilities outside [0, 1].")
  }

  invisible(TRUE)
}

for (scenario_index in seq_along(scenarios)) {
  validate_scenario(scenarios[[scenario_index]], scenario_index, comparison_settings$dose_levels)
}

scenario_true_utility_table <- bind_rows(lapply(seq_along(scenarios), function(scenario_index) {
  scenario <- scenarios[[scenario_index]]
  true_utilities <- calculate_true_utilities(
    scenario,
    comparison_settings$dose_levels,
    utility_table
  )
  max_utility <- max(true_utilities)

  data.frame(
    scenario = scenario$name,
    dose = paste("Dose", comparison_settings$dose_levels),
    dose_level = comparison_settings$dose_levels,
    true_utility = true_utilities,
    is_true_optimal = abs(true_utilities - max_utility) < 1e-10,
    stringsAsFactors = FALSE
  )
}))

scenario_optimal_table <- scenario_true_utility_table %>%
  filter(is_true_optimal) %>%
  group_by(scenario) %>%
  summarise(
    true_optimal_dose = paste(dose, collapse = ", "),
    true_optimal_utility = first(true_utility),
    .groups = "drop"
  )

scenario_true_utility_wide <- scenario_true_utility_table %>%
  mutate(utility_label = paste0("true_utility_Dose_", dose_level)) %>%
  select(scenario, utility_label, true_utility) %>%
  tidyr::pivot_wider(
    names_from = utility_label,
    values_from = true_utility
  )

scenario_truth_table <- bind_rows(lapply(seq_along(scenarios), function(scenario_index) {
  scenario <- scenarios[[scenario_index]]
  marginal_toxicity <- calculate_marginal(scenario$p_YT_given_I, scenario$p_YI)
  marginal_efficacy <- calculate_marginal(scenario$p_YE_given_I, scenario$p_YI)

  data.frame(
    scenario = scenario$name,
    p_YI = format_prob_vector(scenario$p_YI),
    marginal_p_T = format_prob_vector(marginal_toxicity),
    marginal_p_E = format_prob_vector(marginal_efficacy),
    rho0 = scenario$rho0,
    rho1 = scenario$rho1,
    stringsAsFactors = FALSE
  )
})) %>%
  left_join(scenario_optimal_table, by = "scenario") %>%
  left_join(scenario_true_utility_wide, by = "scenario")

scenario_truth_display <- scenario_truth_table
utility_columns <- grep("^true_utility_Dose_", names(scenario_truth_display), value = TRUE)
scenario_truth_display[utility_columns] <- lapply(
  scenario_truth_display[utility_columns],
  function(values) round(values, 1)
)
scenario_truth_display$true_optimal_utility <- round(scenario_truth_display$true_optimal_utility, 1)

kable(scenario_truth_display, caption = "Scenario truth values and true optimal dose")
```

## Run Scenario Simulations

This chunk runs the backend simulations automatically.

```{r run_scenarios}
#| echo: false
n_simulations <- comparison_settings$n_simulations
progress_every <- if (quick_mode) 1 else max(1, floor(n_simulations / 20))
scenario_seed_stride <- max(100000, n_simulations + 1)

get_result_scalar <- function(result, name, default) {
  value <- result[[name]]
  if (is.null(value) || length(value) == 0) {
    return(default)
  }
  value[[1]]
}

run_one_scenario <- function(scenario, scenario_index) {
  cat("\nRunning scenario ", scenario_index, "/", length(scenarios), ": ", scenario$name, "\n", sep = "")
  scenario_start_time <- Sys.time()
  scenario_seed <- comparison_settings$seed + (scenario_index - 1) * scenario_seed_stride

  scenario_results <- lapply(seq_len(n_simulations), function(sim_id) {
    if (sim_id == 1 || sim_id %% progress_every == 0 || sim_id == n_simulations) {
      cat("  Simulation ", sim_id, " of ", n_simulations, "\n", sep = "")
    }
    run_trial_simulation(
      comparison_trial_config,
      scenario$p_YI,
      scenario$p_YT_given_I,
      scenario$p_YE_given_I,
      scenario$rho0,
      scenario$rho1,
      seed = scenario_seed + sim_id - 1
    )
  })

  scenario_duration <- difftime(Sys.time(), scenario_start_time, units = "mins")
  cat("Completed scenario ", scenario$name, " in ", round(scenario_duration, 2), " minutes.\n", sep = "")

  scenario_metrics <- data.frame(
    scenario = scenario$name,
    simulation = seq_along(scenario_results),
    terminated_early = vapply(scenario_results, function(result) {
      isTRUE(get_result_scalar(result, "terminated_early", FALSE))
    }, logical(1)),
    termination_stage = vapply(scenario_results, function(result) {
      as.integer(get_result_scalar(result, "termination_stage", NA_integer_))
    }, integer(1)),
    final_od = vapply(scenario_results, function(result) {
      as.integer(get_result_scalar(result, "final_od", NA_integer_))
    }, integer(1)),
    poc_validated = vapply(scenario_results, function(result) {
      isTRUE(get_result_scalar(result, "poc_validated", FALSE))
    }, logical(1)),
    poc_probability = vapply(scenario_results, function(result) {
      as.numeric(get_result_scalar(result, "poc_probability", NA_real_))
    }, numeric(1)),
    final_utility = vapply(scenario_results, function(result) {
      as.numeric(get_result_scalar(result, "final_utility", NA_real_))
    }, numeric(1)),
    n_participants = vapply(scenario_results, function(result) {
      nrow(result$all_data)
    }, integer(1)),
    stringsAsFactors = FALSE
  )

  scenario_metrics$selection <- ifelse(
    is.na(scenario_metrics$final_od),
    "No OD",
    paste("Dose", scenario_metrics$final_od)
  )

  scenario_metrics
}

scenario_metrics <- bind_rows(lapply(seq_along(scenarios), function(scenario_index) {
  run_one_scenario(scenarios[[scenario_index]], scenario_index)
}))
```

## Results

This chunk organizes the scenario outputs into final tables.

```{r results}
#| echo: false
scenario_output_dir <- file.path(project_root_abs, "results", "scenario_comparison")
scenario_output_path <- function(filename) file.path(scenario_output_dir, filename)

dir.create(scenario_output_dir, recursive = TRUE, showWarnings = FALSE)

selection_levels <- c(paste("Dose", comparison_settings$dose_levels), "No OD")

scenario_selection_table <- scenario_metrics %>%
  mutate(selection = factor(selection, levels = selection_levels)) %>%
  count(scenario, selection, name = "n") %>%
  tidyr::complete(scenario, selection, fill = list(n = 0)) %>%
  group_by(scenario) %>%
  mutate(selection_rate = n / sum(n)) %>%
  ungroup() %>%
  mutate(
    selection_label = ifelse(
      as.character(selection) == "No OD",
      "selection_rate_No_OD",
      paste0("selection_rate_Dose_", sub("^Dose ", "", as.character(selection)))
    )
  )

scenario_selection_wide <- scenario_selection_table %>%
  select(scenario, selection_label, selection_rate) %>%
  tidyr::pivot_wider(
    names_from = selection_label,
    values_from = selection_rate,
    values_fill = 0
  )

true_optimal_selection_table <- scenario_selection_table %>%
  mutate(selection = as.character(selection)) %>%
  inner_join(
    scenario_true_utility_table %>%
      filter(is_true_optimal) %>%
      select(scenario, selection = dose),
    by = c("scenario", "selection")
  ) %>%
  group_by(scenario) %>%
  summarise(
    selection_rate_true_optimal = sum(selection_rate),
    .groups = "drop"
  )

scenario_summary_table <- scenario_metrics %>%
  group_by(scenario) %>%
  summarise(
    n_simulations = n(),
    mean_participants = mean(n_participants),
    early_stop_rate = mean(terminated_early),
    poc_validation_rate = mean(poc_validated),
    no_od_rate = mean(is.na(final_od)),
    mean_poc_probability = mean(poc_probability, na.rm = TRUE),
    mean_final_utility = mean(final_utility, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(scenario_optimal_table, by = "scenario") %>%
  left_join(true_optimal_selection_table, by = "scenario") %>%
  left_join(scenario_selection_wide, by = "scenario") %>%
  relocate(true_optimal_dose, true_optimal_utility, selection_rate_true_optimal, .after = scenario) %>%
  arrange(scenario)

display_rate_columns <- grep("rate", names(scenario_summary_table), value = TRUE)
scenario_summary_display <- scenario_summary_table
scenario_summary_display[display_rate_columns] <- lapply(
  scenario_summary_display[display_rate_columns],
  function(values) sprintf("%.1f%%", 100 * values)
)
scenario_summary_display$mean_participants <- round(scenario_summary_display$mean_participants, 1)
scenario_summary_display$mean_poc_probability <- round(scenario_summary_display$mean_poc_probability, 3)
scenario_summary_display$mean_final_utility <- round(scenario_summary_display$mean_final_utility, 1)
scenario_summary_display$true_optimal_utility <- round(scenario_summary_display$true_optimal_utility, 1)

kable(scenario_summary_display, caption = "Scenario comparison summary with true optimal dose")

write.csv(scenario_truth_table, scenario_output_path("scenario_truth_table.csv"), row.names = FALSE)
write.csv(scenario_true_utility_table, scenario_output_path("scenario_true_utility_table.csv"), row.names = FALSE)
write.csv(scenario_metrics, scenario_output_path("scenario_metrics.csv"), row.names = FALSE)
write.csv(scenario_selection_table, scenario_output_path("scenario_selection_table.csv"), row.names = FALSE)
write.csv(scenario_summary_table, scenario_output_path("scenario_summary_table.csv"), row.names = FALSE)

cat("\nSaved scenario truth table to ", scenario_output_path("scenario_truth_table.csv"), "\n", sep = "")
cat("Saved true utility table to ", scenario_output_path("scenario_true_utility_table.csv"), "\n", sep = "")
cat("Saved simulation-level metrics to ", scenario_output_path("scenario_metrics.csv"), "\n", sep = "")
cat("Saved selection table to ", scenario_output_path("scenario_selection_table.csv"), "\n", sep = "")
cat("Saved final summary table to ", scenario_output_path("scenario_summary_table.csv"), "\n", sep = "")
```

## Summary

This notebook is for comparing scenario-level operating characteristics. For one detailed scenario with plots, use `simulation_notebook.qmd`. For calibration, use `threshold_calibration_notebook.qmd` and `poc_calibration_notebook.qmd`.
