---
title: "Bayesian Dose-Finding Trial Simulation"
format:
  html:
    toc: true
    code-fold: true
  pdf: default
execute:
  message: false
  warning: false
editor: visual
---

## Introduction

This notebook is the standard user interface for running Bayesian dose-finding trial simulations. Edit only the **User Settings** chunk, then click **Run All** or **Render** in RStudio. The setup and backend simulation chunks run automatically.

Use quick mode for a fast smoke test with 5 trial simulations. Switch `quick_mode` to `FALSE` for the production default of 2,000 trial simulations.

## User Settings

Edit this chunk only for routine use.

```{r user_settings}
quick_mode <- FALSE
dose_levels <- c(1, 2, 3, 4, 5)

simulation_settings <- list(
  dose_levels = dose_levels,
  n_stages = if (quick_mode) 3 else 5,
  n_simulations = if (quick_mode) 5 else 2000,
  cohort_size = 15,
  phi_T = 0.30,
  c_T = 0.35,
  phi_E = 0.25,
  c_E = 0.55,
  phi_I = 0.20,
  c_I = 0.50,
  c_poc = 0.90,
  delta_poc = 0.8,
  use_calibration_results = TRUE,
  threshold_calibration_results_path = "results/threshold_calibration/threshold_calibration_results.rds",
  poc_calibration_results_path = "results/notebook_calibration/poc_calibration_results.rds",
  seed = 11118,
  p_YI = c(0.10, 0.30, 0.50, 0.60, 0.70),
  p_YT_given_I = matrix(c(
    0.05, 0.10, 0.12, 0.18, 0.25,
    0.08, 0.12, 0.15, 0.25, 0.35
  ), nrow = length(dose_levels), ncol = 2),
  p_YE_given_I = matrix(c(
    0.10, 0.20, 0.35, 0.45, 0.50,
    0.30, 0.50, 0.70, 0.80, 0.75
  ), nrow = length(dose_levels), ncol = 2),
  rho0 = 1.5,
  rho1 = 2
)
```

## Internal Setup

This chunk loads the backend code automatically.

```{r setup}
#| echo: false
library(knitr)
library(ggplot2)

# Find and use the project root
project_root_candidates <- c(".", "..", "../..")
project_root_matches <- project_root_candidates[
  file.exists(file.path(project_root_candidates, "DoseFinding.Rproj"))
]
if (length(project_root_matches) == 0) {
  stop("Could not find project root containing DoseFinding.Rproj.")
}
project_root <- normalizePath(project_root_matches[[1]], winslash = "/", mustWork = TRUE)
project_root_abs <- project_root
setwd(project_root)
cat("Project root:", project_root, "\n")

# Source core files
source(file.path(project_root, "src/core/config.R"))
source(file.path(project_root, "src/utils/helpers.R"))
source(file.path(project_root, "src/utils/plotting_extensions.R"))
source(file.path(project_root, "src/core/simulate_data.R"))
source(file.path(project_root, "src/core/model_utils.R"))
source(file.path(project_root, "src/decision/dose_decision.R"))
source(file.path(project_root, "src/core/main.R"))

resolve_project_path <- function(path) {
  if (is.null(path) || length(path) == 0 || is.na(path) || path == "") {
    return(path)
  }
  if (grepl("^(/|[A-Za-z]:[\\\\/])", path)) {
    return(path)
  }
  file.path(project_root_abs, path)
}

simulation_settings$threshold_calibration_results_path <- resolve_project_path(
  simulation_settings$threshold_calibration_results_path
)
simulation_settings$poc_calibration_results_path <- resolve_project_path(
  simulation_settings$poc_calibration_results_path
)
```

## Calibrated Inputs

This chunk reads saved calibration results when available.

```{r calibrated_inputs}
#| echo: false
calibration_source <- c(
  c_T = "User Settings fallback",
  c_I = "User Settings fallback",
  c_E = "User Settings fallback",
  c_poc = "User Settings fallback"
)

apply_scalar_calibration_value <- function(settings, sources, name, value, source_label) {
  if (!is.numeric(value) || length(value) != 1 || is.na(value)) {
    stop("Invalid calibration value for ", name, ".")
  }
  settings[[name]] <- as.numeric(value)
  sources[[name]] <- source_label
  list(settings = settings, sources = sources)
}

if (isTRUE(simulation_settings$use_calibration_results)) {
  threshold_path <- simulation_settings$threshold_calibration_results_path
  if (file.exists(threshold_path)) {
    threshold_results <- readRDS(threshold_path)
    recommended_thresholds <- threshold_results$recommended_thresholds
    required_thresholds <- c("c_T", "c_I", "c_E")
    missing_thresholds <- setdiff(required_thresholds, names(recommended_thresholds))
    if (length(missing_thresholds) > 0) {
      stop(
        "Threshold calibration results are missing: ",
        paste(missing_thresholds, collapse = ", ")
      )
    }
    for (name in required_thresholds) {
      applied <- apply_scalar_calibration_value(
        simulation_settings,
        calibration_source,
        name,
        recommended_thresholds[[name]],
        paste0("Threshold calibration results: ", threshold_path)
      )
      simulation_settings <- applied$settings
      calibration_source <- applied$sources
    }
  } else {
    warning(
      "Threshold calibration results were not found at ",
      threshold_path,
      ". Using c_T, c_I, and c_E from User Settings."
    )
  }

  poc_path <- simulation_settings$poc_calibration_results_path
  if (file.exists(poc_path)) {
    poc_results <- readRDS(poc_path)
    applied <- apply_scalar_calibration_value(
      simulation_settings,
      calibration_source,
      "c_poc",
      poc_results$optimal_c_poc,
      paste0("PoC calibration results: ", poc_path)
    )
    simulation_settings <- applied$settings
    calibration_source <- applied$sources
  } else {
    warning(
      "PoC calibration results were not found at ",
      poc_path,
      ". Using c_poc from User Settings."
    )
  }
}

calibration_input_table <- data.frame(
  parameter = names(calibration_source),
  current_value = vapply(names(calibration_source), function(name) {
    simulation_settings[[name]]
  }, numeric(1)),
  source = unname(calibration_source),
  stringsAsFactors = FALSE
)

kable(calibration_input_table, caption = "Calibration values used in this simulation")
```

## Internal Configuration

This chunk converts the user settings into the backend trial configuration.

```{r config}
#| echo: false
trial_config <- list(
  dose_levels = simulation_settings$dose_levels,
  n_stages = simulation_settings$n_stages,
  cohort_size = simulation_settings$cohort_size,
  phi_T = simulation_settings$phi_T,
  c_T = simulation_settings$c_T,
  phi_E = simulation_settings$phi_E,
  c_E = simulation_settings$c_E,
  phi_I = simulation_settings$phi_I,
  c_I = simulation_settings$c_I,
  c_poc = simulation_settings$c_poc,
  delta_poc = simulation_settings$delta_poc,
  enable_early_termination = TRUE,
  log_early_termination = FALSE,
  verbose_logging = FALSE
)

p_YI <- simulation_settings$p_YI
p_YT_given_I <- simulation_settings$p_YT_given_I
p_YE_given_I <- simulation_settings$p_YE_given_I
rho0 <- simulation_settings$rho0
rho1 <- simulation_settings$rho1

utility_table <- array(0, dim = c(2, 2, 2), dimnames = list(
  E = c(0, 1),
  T = c(0, 1),
  I = c(0, 1)
))

utility_table[1, 1, 1] <- 0   # E=0, T=0, I=0
utility_table[2, 1, 1] <- 80  # E=1, T=0, I=0
utility_table[1, 2, 1] <- 0   # E=0, T=1, I=0
utility_table[2, 2, 1] <- 30  # E=1, T=1, I=0

utility_table[1, 1, 2] <- 10  # E=0, T=0, I=1
utility_table[2, 1, 2] <- 100 # E=1, T=0, I=1
utility_table[1, 2, 2] <- 0   # E=0, T=1, I=1
utility_table[2, 2, 2] <- 40  # E=1, T=1, I=1

trial_config$utility_table <- utility_table
```

## Simulation

This chunk runs the backend simulation automatically. It repeats the full trial `n_simulations` times. The workflow inside each replicate is:

1.  **Stage 1**: Equal randomization to all dose levels
2.  **Interim Analysis**: Update admissible set based on posterior probabilities
3.  **Early Termination Check**: Terminate if admissible set is empty (checked after interim analysis, before adaptive randomization)
4.  **Adaptive Randomization**: Allocate patients based on utility scores (Stages 2+ only, only if trial continues)
5.  **Final Selection**: Choose OD with highest utility from admissible set + PoC validation

```{r simulation}
#| echo: false
n_simulations <- simulation_settings$n_simulations
progress_every <- if (quick_mode) 1 else max(1, floor(n_simulations / 20))

cat("Running", n_simulations, "trial simulations...\n")
simulation_start_time <- Sys.time()

simulation_results <- lapply(seq_len(n_simulations), function(sim_id) {
  if (sim_id == 1 || sim_id %% progress_every == 0 || sim_id == n_simulations) {
    cat("  Simulation", sim_id, "of", n_simulations, "\n")
  }
  run_trial_simulation(
    trial_config,
    p_YI,
    p_YT_given_I,
    p_YE_given_I,
    rho0,
    rho1,
    seed = simulation_settings$seed + sim_id - 1
  )
})

simulation_duration <- difftime(Sys.time(), simulation_start_time, units = "mins")
cat("Completed", n_simulations, "trial simulations in", round(simulation_duration, 2), "minutes.\n")

get_result_scalar <- function(result, name, default) {
  value <- result[[name]]
  if (is.null(value) || length(value) == 0) {
    return(default)
  }
  value[[1]]
}

simulation_metrics <- data.frame(
  simulation = seq_along(simulation_results),
  terminated_early = vapply(simulation_results, function(result) {
    isTRUE(get_result_scalar(result, "terminated_early", FALSE))
  }, logical(1)),
  termination_stage = vapply(simulation_results, function(result) {
    as.integer(get_result_scalar(result, "termination_stage", NA_integer_))
  }, integer(1)),
  final_od = vapply(simulation_results, function(result) {
    as.integer(get_result_scalar(result, "final_od", NA_integer_))
  }, integer(1)),
  poc_validated = vapply(simulation_results, function(result) {
    isTRUE(get_result_scalar(result, "poc_validated", FALSE))
  }, logical(1)),
  poc_probability = vapply(simulation_results, function(result) {
    as.numeric(get_result_scalar(result, "poc_probability", NA_real_))
  }, numeric(1)),
  final_utility = vapply(simulation_results, function(result) {
    as.numeric(get_result_scalar(result, "final_utility", NA_real_))
  }, numeric(1)),
  n_participants = vapply(simulation_results, function(result) {
    nrow(result$all_data)
  }, integer(1)),
  stringsAsFactors = FALSE
)

simulation_metrics$selection <- ifelse(
  is.na(simulation_metrics$final_od),
  "No OD",
  paste("Dose", simulation_metrics$final_od)
)

true_toxicity <- p_YT_given_I[, 1] * (1 - p_YI) + p_YT_given_I[, 2] * p_YI
true_efficacy <- p_YE_given_I[, 1] * (1 - p_YI) + p_YE_given_I[, 2] * p_YI
overdose_doses <- trial_config$dose_levels[true_toxicity > trial_config$phi_T]
simulation_metrics$selected_overdose <- !is.na(simulation_metrics$final_od) &
  simulation_metrics$final_od %in% overdose_doses
```

## Results

This chunk prints and plots Monte Carlo summary results across all simulations. All plots in this section are aggregate summaries, not first-simulation examples.

```{r results}
#| echo: false
plot_output_dir <- file.path(project_root_abs, "results", "plots")
simulation_output_dir <- file.path(project_root_abs, "results", "simulation")
plot_output_path <- function(filename) file.path(plot_output_dir, filename)
simulation_output_path <- function(filename) file.path(simulation_output_dir, filename)

dir.create(plot_output_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(simulation_output_dir, recursive = TRUE, showWarnings = FALSE)

# Print final results
cat("
--- Monte Carlo Summary ---
")

cat("Number of simulations:", nrow(simulation_metrics), "\n")
cat("Early termination rate:", sprintf("%.1f%%", 100 * mean(simulation_metrics$terminated_early)), "\n")
cat("PoC validation rate:", sprintf("%.1f%%", 100 * mean(simulation_metrics$poc_validated)), "\n")
cat("Mean participants:", round(mean(simulation_metrics$n_participants), 1), "\n")

selection_summary <- simulation_metrics %>%
  count(selection, name = "n") %>%
  mutate(rate = n / nrow(simulation_metrics)) %>%
  arrange(desc(n))

kable(selection_summary, caption = "Final selection summary across simulations")

dose_truth_table <- data.frame(
  dose = trial_config$dose_levels,
  true_marginal_toxicity = true_toxicity,
  overdose = true_toxicity > trial_config$phi_T,
  stringsAsFactors = FALSE
)

dose_selection_summary <- data.frame(
  dose = trial_config$dose_levels,
  stringsAsFactors = FALSE
) %>%
  left_join(
    simulation_metrics %>%
      filter(!is.na(final_od)) %>%
      count(final_od, name = "n_selected"),
    by = c("dose" = "final_od")
  ) %>%
  mutate(
    n_selected = dplyr::coalesce(n_selected, 0L),
    selection_probability = n_selected / nrow(simulation_metrics),
    true_marginal_toxicity = true_toxicity[match(dose, trial_config$dose_levels)],
    overdose = true_marginal_toxicity > trial_config$phi_T
  )

overdose_selection_summary <- data.frame(
  n_simulations = nrow(simulation_metrics),
  n_selected_overdose = sum(simulation_metrics$selected_overdose),
  overdose_selection_probability = mean(simulation_metrics$selected_overdose),
  phi_T = trial_config$phi_T,
  overdose_doses = ifelse(length(overdose_doses) > 0, paste(overdose_doses, collapse = ", "), "None"),
  stringsAsFactors = FALSE
)

kable(dose_truth_table, caption = "True marginal toxicity by dose")
kable(dose_selection_summary, caption = "Probability of selecting each dose as ODB")
kable(overdose_selection_summary, caption = "Probability of selecting an overdose as ODB")

cat(
  "Probability of selecting an overdose as ODB:",
  sprintf("%.1f%%", 100 * overdose_selection_summary$overdose_selection_probability),
  "\n"
)

write.csv(simulation_metrics, simulation_output_path("simulation_metrics.csv"), row.names = FALSE)

allocation_probability_summary <- bind_rows(lapply(seq_along(simulation_results), function(sim_id) {
  allocation_probabilities <- simulation_results[[sim_id]]$all_alloc_probs
  if (is.null(allocation_probabilities) || nrow(allocation_probabilities) == 0) {
    return(data.frame())
  }
  allocation_probabilities$simulation <- sim_id
  allocation_probabilities
})) %>%
  mutate(
    Stage = as.integer(Stage),
    Dose = factor(Dose, levels = trial_config$dose_levels)
  ) %>%
  group_by(Stage, Dose) %>%
  summarise(
    mean_prob = mean(Prob),
    n_active_trials = n(),
    .groups = "drop"
  )

participant_allocation_by_sim <- bind_rows(lapply(seq_along(simulation_results), function(sim_id) {
  participant_counts <- simulation_results[[sim_id]]$all_data %>%
    mutate(
      d = as.integer(d),
      stage = as.integer(stage)
    ) %>%
    group_by(d, stage) %>%
    summarise(n_participants = n(), .groups = "drop")

  participant_allocation_grid <- expand.grid(
    simulation = sim_id,
    d = trial_config$dose_levels,
    stage = seq_len(trial_config$n_stages),
    KEEP.OUT.ATTRS = FALSE
  )

  participant_allocation_grid %>%
    left_join(participant_counts, by = c("d", "stage")) %>%
    mutate(n_participants = dplyr::coalesce(n_participants, 0L))
}))

stage_enrollment_summary <- participant_allocation_by_sim %>%
  group_by(simulation, stage) %>%
  summarise(stage_participants = sum(n_participants), .groups = "drop") %>%
  group_by(stage) %>%
  summarise(
    planned_simulations = n(),
    active_trials = sum(stage_participants > 0),
    active_trial_rate = active_trials / planned_simulations,
    unconditional_mean_participants = mean(stage_participants),
    conditional_mean_participants = ifelse(
      active_trials > 0,
      mean(stage_participants[stage_participants > 0]),
      NA_real_
    ),
    .groups = "drop"
  ) %>%
  mutate(
    stage_num = as.integer(stage),
    stage = factor(
      stage_num,
      levels = seq_len(trial_config$n_stages),
      labels = paste("Stage", seq_len(trial_config$n_stages))
    )
  )

participant_allocation_summary <- participant_allocation_by_sim %>%
  group_by(d, stage) %>%
  summarise(
    mean_participants = mean(n_participants),
    total_participants = sum(n_participants),
    .groups = "drop"
  ) %>%
  mutate(
    d = factor(d, levels = trial_config$dose_levels),
    stage_num = as.integer(stage),
    stage = factor(
      stage_num,
      levels = seq_len(trial_config$n_stages),
      labels = paste("Stage", seq_len(trial_config$n_stages))
    )
  ) %>%
  arrange(d, stage_num)

final_participant_allocation_summary <- participant_allocation_by_sim %>%
  group_by(simulation, d) %>%
  summarise(final_participants = sum(n_participants), .groups = "drop") %>%
  group_by(d) %>%
  summarise(
    mean_final_participants = mean(final_participants),
    sd_final_participants = sd(final_participants),
    total_participants = sum(final_participants),
    .groups = "drop"
  ) %>%
  mutate(d = factor(d, levels = trial_config$dose_levels)) %>%
  arrange(d)

write.csv(
  allocation_probability_summary,
  simulation_output_path("allocation_probability_summary.csv"),
  row.names = FALSE
)

write.csv(
  participant_allocation_summary,
  simulation_output_path("participant_allocation_summary.csv"),
  row.names = FALSE
)

write.csv(
  final_participant_allocation_summary,
  simulation_output_path("final_participant_allocation_summary.csv"),
  row.names = FALSE
)

write.csv(
  dose_selection_summary,
  simulation_output_path("dose_selection_summary.csv"),
  row.names = FALSE
)

write.csv(
  overdose_selection_summary,
  simulation_output_path("overdose_selection_summary.csv"),
  row.names = FALSE
)

write.csv(
  stage_enrollment_summary,
  simulation_output_path("stage_enrollment_summary.csv"),
  row.names = FALSE
)

cat("\nSaved simulation-level metrics to results/simulation/simulation_metrics.csv\n")
cat("Saved mean allocation probabilities to results/simulation/allocation_probability_summary.csv\n")
cat("Saved mean participant allocation to results/simulation/participant_allocation_summary.csv\n")
cat("Saved mean final participant allocation to results/simulation/final_participant_allocation_summary.csv\n")
cat("Saved dose selection probabilities to results/simulation/dose_selection_summary.csv\n")
cat("Saved overdose selection summary to results/simulation/overdose_selection_summary.csv\n")
cat("Saved stage enrollment summary to results/simulation/stage_enrollment_summary.csv\n")

p_selection <- ggplot(selection_summary, aes(x = reorder(selection, -n), y = rate, fill = selection)) +
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels = function(x) paste0(round(100 * x), "%")) +
  labs(
    title = "Final Selection Rate Across Simulations",
    x = "Selection",
    y = "Selection Rate"
  ) +
  theme_bw(base_size = 14) +
  theme(panel.grid = element_blank())
print(p_selection)

p_dose_selection <- ggplot(
  dose_selection_summary,
  aes(x = factor(dose, levels = trial_config$dose_levels), y = selection_probability, fill = overdose)
) +
  geom_col(width = 0.7, color = "black") +
  geom_text(
    aes(label = paste0(round(100 * selection_probability, 1), "%")),
    vjust = -0.3,
    size = 3.8
  ) +
  scale_y_continuous(labels = function(x) paste0(round(100 * x), "%"), limits = c(0, 1)) +
  scale_fill_manual(values = c("FALSE" = "#56B4E9", "TRUE" = "#D55E00"), name = "True overdose") +
  labs(
    title = "Probability of Selecting Each Dose as ODB",
    subtitle = paste("True overdose means marginal P(T) >", trial_config$phi_T),
    x = "Dose Level",
    y = "Selection Probability"
  ) +
  theme_bw(base_size = 14) +
  theme(panel.grid = element_blank(), plot.title = element_text(hjust = 0.5))
print(p_dose_selection)

p_overdose_selection <- ggplot(
  overdose_selection_summary,
  aes(x = "Selected overdose", y = overdose_selection_probability)
) +
  geom_col(width = 0.45, fill = "#D55E00", color = "black") +
  geom_text(
    aes(label = paste0(round(100 * overdose_selection_probability, 1), "%")),
    vjust = -0.4,
    size = 4
  ) +
  scale_y_continuous(labels = function(x) paste0(round(100 * x), "%"), limits = c(0, 1)) +
  labs(
    title = "Probability of Selecting an Overdose as ODB",
    subtitle = paste("Overdose doses:", overdose_selection_summary$overdose_doses),
    x = NULL,
    y = "Probability"
  ) +
  theme_bw(base_size = 14) +
  theme(panel.grid = element_blank(), plot.title = element_text(hjust = 0.5))
print(p_overdose_selection)

overlap_position <- position_dodge(width = 0.22)
p_alloc_time <- ggplot(
  allocation_probability_summary,
  aes(x = Stage, y = mean_prob, color = Dose, group = Dose)
) +
  geom_line(linewidth = 1, position = overlap_position) +
  geom_point(size = 3, position = overlap_position) +
  labs(
    title = "Mean Allocation Probabilities Across Simulations",
    subtitle = paste("Averaged among active trials at each stage; simulations:", nrow(simulation_metrics)),
    x = "Stage",
    y = "Mean Allocation Probability",
    color = "Dose Level"
  ) +
  scale_color_manual(values = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#CC79A7")) +
  scale_x_continuous(breaks = seq_len(trial_config$n_stages)) +
  theme_bw(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5),
    axis.line = element_line(color = "black")
  )
print(p_alloc_time)

aggregate_posterior_summary <- function(summary_name, group_vars) {
  posterior_rows <- bind_rows(lapply(seq_along(simulation_results), function(sim_id) {
    posterior_summary <- simulation_results[[sim_id]]$posterior_summaries[[summary_name]]
    if (is.null(posterior_summary) || nrow(posterior_summary) == 0) {
      return(data.frame())
    }
    posterior_summary$simulation <- sim_id
    posterior_summary
  }))

  posterior_rows %>%
    group_by(across(all_of(group_vars))) %>%
    summarise(pava_mean = mean(pava_mean, na.rm = TRUE), .groups = "drop")
}

mean_immune_summary <- aggregate_posterior_summary("imm", "d")
mean_toxicity_summary <- aggregate_posterior_summary("tox", c("d", "Y_I"))
mean_efficacy_summary <- aggregate_posterior_summary("eff", c("d", "Y_I"))

plot_posterior_summary(
  mean_immune_summary,
  title = "Mean Final Immune Response Posterior Across Simulations",
  file_path = plot_output_path("mean_immune_response.png"),
  style = "modern"
)
plot_posterior_summary(
  mean_toxicity_summary,
  title = "Mean Final Toxicity Posterior Across Simulations",
  group_col = "Y_I",
  file_path = plot_output_path("mean_toxicity.png"),
  style = "modern"
)
plot_posterior_summary(
  mean_efficacy_summary,
  title = "Mean Final Efficacy Posterior Across Simulations",
  group_col = "Y_I",
  file_path = plot_output_path("mean_efficacy.png"),
  style = "modern"
)

# Create dose-response curves similar to reference code
cat("\n=== Creating Dose-Response Curves ===\n")

# Calculate true utilities for all dose levels using true probabilities
true_utility <- sapply(1:length(trial_config$dose_levels), 
                       calculate_utility_from_true_probs, 
                       p_YI = p_YI, 
                       p_YT_given_I = p_YT_given_I, 
                       p_YE_given_I = p_YE_given_I,
                       utility_table = utility_table)

# Create dose-response curves plot
dose_response_plot <- plot_dose_response_curves(
  toxicity_data = true_toxicity,
  efficacy_data = true_efficacy,
  utility_data = true_utility,
  title = "True Dose-Response Curves",
  file_path = plot_output_path("dose_response_curves.png")
)
print(dose_response_plot)

# Plot 1: Unconditional mean allocation by dose level and stage
p_alloc <- ggplot(participant_allocation_summary, aes(x = d, y = mean_participants, fill = stage)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "black") +
  labs(
    title = "Unconditional Mean Participant Allocation by Dose Level and Stage",
    x = "Dose Level",
    y = "Mean Participants per Planned Simulation",
    subtitle = "Stages after early termination count as 0 for that simulation"
  ) +
  scale_fill_manual(name = "Stage", values = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#CC79A7")) +
  theme_bw(base_size = 16) +
  theme(panel.grid = element_blank(),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        axis.line = element_line(color = "black"))
print(p_alloc)

kable(
  final_participant_allocation_summary,
  caption = "Final mean participant allocation by dose across simulations"
)

p_final_alloc <- ggplot(final_participant_allocation_summary, aes(x = d, y = mean_final_participants)) +
  geom_col(width = 0.7, fill = "#009E73", color = "black") +
  geom_text(
    aes(label = round(mean_final_participants, 1)),
    vjust = -0.3,
    size = 3.8
  ) +
  labs(
    title = "Final Mean Participants by Dose",
    subtitle = "Averaged across all planned simulations after early termination",
    x = "Dose Level",
    y = "Mean Participants"
  ) +
  theme_bw(base_size = 14) +
  theme(panel.grid = element_blank(), plot.title = element_text(hjust = 0.5))
print(p_final_alloc)

# Plot 2: Cumulative allocation over stages
cumulative_summary <- participant_allocation_summary %>%
  arrange(d, stage_num) %>%
  group_by(d) %>%
  mutate(mean_cumulative_participants = cumsum(mean_participants)) %>%
  ungroup()

p_cumulative <- ggplot(cumulative_summary, aes(x = stage, y = mean_cumulative_participants, color = d, group = d)) +
  geom_line(linewidth = 1.5, position = overlap_position) +
  geom_point(size = 3, position = overlap_position) +
  labs(
    title = "Unconditional Mean Cumulative Participant Allocation Over Stages",
    x = "Stage",
    y = "Mean Cumulative Participants per Planned Simulation",
    color = "Dose Level",
    subtitle = paste("Mean final total participants:", round(mean(simulation_metrics$n_participants), 1))
  ) +
  scale_color_manual(values = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#CC79A7")) +
  theme_bw(base_size = 16) +
  theme(panel.grid = element_blank(),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        axis.line = element_line(color = "black"))
print(p_cumulative)

# Print summary statistics
cat("\n=== MEAN PARTICIPANT ALLOCATION SUMMARY ===\n")
cat("Mean final total participants:", round(mean(simulation_metrics$n_participants), 1), "\n")
cat("Unconditional mean participants per planned stage:\n")
for(i in 1:nrow(stage_enrollment_summary)) {
  cat(
    "  ", as.character(stage_enrollment_summary$stage[i]), ": ",
    round(stage_enrollment_summary$unconditional_mean_participants[i], 1),
    " mean participants; ",
    stage_enrollment_summary$active_trials[i], "/",
    stage_enrollment_summary$planned_simulations[i],
    " trials reached this stage; conditional mean among reached trials = ",
    round(stage_enrollment_summary$conditional_mean_participants[i], 1),
    "\n",
    sep = ""
  )
}

```

## Summary

This notebook runs repeated Bayesian dose-finding trial simulations and reports aggregate Monte Carlo outputs. The plots include:

- **Selection rates** across all simulations
- **ODB selection probability by dose** across all simulations
- **Overdose-as-ODB selection probability** using true marginal toxicity above `phi_T`
- **Mean allocation probabilities** across all simulations
- **Mean final posterior summaries** across all simulations
- **Mean participant allocation** across all simulations
- **Final mean participants by dose** across all simulations
- **True dose-response curves** with toxicity, efficacy, and utility

For threshold calibration and PoC calibration, see the dedicated `threshold_calibration_notebook.qmd` and `poc_calibration_notebook.qmd`.

All plots in this notebook are aggregate summaries from the simulation run.
