---
title: "Threshold Calibration Notebook"
subtitle: "Separate calibration of c_T, c_I, and c_E before PoC calibration"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show Code"
    embed-resources: true
    self-contained: true
execute:
  message: false
  warning: false
  echo: false
editor: visual
---

## 1. Purpose

This notebook calibrates the three admissibility credibility cutoffs separately:

1.  `c_T`: evidence required to keep a dose when toxicity is judged acceptable.
2.  `c_I`: evidence required to keep a dose when immune response is judged acceptable.
3.  `c_E`: evidence required to keep a dose when efficacy is judged acceptable.

The calibration target is the final admissible set missing rate: the proportion of simulated trials where the final admissible set is empty under an endpoint-specific unfavorable scenario. The default target is 80%-90%.

After this notebook selects `c_T`, `c_I`, and `c_E`, use those values as fixed inputs in the PoC calibration notebook.

## User Settings

Edit this chunk only for routine use.

```{r user_settings}
#| echo: true

quick_mode <- FALSE

threshold_settings <- list(
  quick_mode = quick_mode,

  dose_levels = c(1, 2, 3, 4, 5),
  n_stages = 5,
  cohort_size = 15,

  phi_T = 0.30,
  phi_E = 0.25,
  phi_I = 0.20,

  c_T_start = 0.55,
  c_E_start = 0.50,
  c_I_start = 0.70,
  c_poc = 0.995,
  delta_poc = 0.8,
  rho0 = 1.5,
  rho1 = 2.0,

  target_missing_range = c(0.80, 0.90),
  c_T_candidates = if (quick_mode) c(0.45, 0.55, 0.65) else seq(0.35, 0.75, by = 0.05),
  c_I_candidates = if (quick_mode) c(0.50, 0.70, 0.90) else seq(0.45, 0.95, by = 0.05),
  c_E_candidates = if (quick_mode) c(0.35, 0.50, 0.65) else seq(0.30, 0.80, by = 0.05),
  n_sim_per_candidate = if (quick_mode) 5 else 500,
  calibration_seed = 11118,

  high_tox_p_I = 0.30,
  high_tox_marginal_p_T = c(0.35, 0.60),
  high_tox_marginal_p_E = 0.40,

  low_immune_p_I = c(0.10, 0.15),
  low_immune_marginal_p_T = 0.15,
  low_immune_marginal_p_E = 0.40,

  low_eff_p_I = 0.30,
  low_eff_marginal_p_T = 0.15,
  low_eff_marginal_p_E = c(0.10, 0.20),

  toxicity_immune_effect = 0,
  efficacy_immune_effect = 0,

  output_dir = "results/threshold_calibration",
  append_history_log = TRUE,
  history_log_path = file.path("results", "threshold_calibration", "threshold_calibration_history.md"),
  run_label = if (quick_mode) "quick smoke test" else "production calibration"
)
```

## 2. Setup

```{r setup}
#| message: false
#| warning: false
#| echo: false

library(dplyr)
library(ggplot2)
library(knitr)

project_root_candidates <- c(".", "..", "../..")
project_root_matches <- project_root_candidates[
  file.exists(file.path(project_root_candidates, "DoseFinding.Rproj"))
]
if (length(project_root_matches) == 0) {
  stop("Could not find project root containing DoseFinding.Rproj.")
}
project_root_abs <- normalizePath(project_root_matches[[1]], winslash = "/", mustWork = TRUE)
setwd(project_root_abs)

source("src/optimization/threshold_calibration.R")

threshold_settings$output_dir <- file.path(project_root_abs, threshold_settings$output_dir)
threshold_settings$history_log_path <- file.path(project_root_abs, threshold_settings$history_log_path)
```

## 3. Parameter Guide

```{r parameter_guide}
#| results: asis

format_setting <- function(value) {
  if (is.logical(value)) {
    return(ifelse(value, "TRUE", "FALSE"))
  }
  if (is.numeric(value)) {
    if (length(value) > 5) {
      diffs <- diff(value)
      if (length(unique(round(diffs, 10))) == 1) {
        step <- diffs[[1]]
        return(paste0(
          round(value[[1]], 4),
          " to ",
          round(value[[length(value)]], 4),
          " by ",
          round(step, 4),
          " (",
          length(value),
          " values)"
        ))
      }
      return(paste0(
        round(value[[1]], 4),
        ", ..., ",
        round(value[[length(value)]], 4),
        " (",
        length(value),
        " values)"
      ))
    }
    return(paste(round(value, 4), collapse = ", "))
  }
  paste(value, collapse = ", ")
}

format_path <- function(path) {
  project_prefix <- paste0(project_root_abs, "/")
  if (startsWith(path, project_prefix)) {
    return(substr(path, nchar(project_prefix) + 1, nchar(path)))
  }
  path
}

escape_html <- function(value) {
  value <- gsub("&", "&amp;", value, fixed = TRUE)
  value <- gsub("<", "&lt;", value, fixed = TRUE)
  value <- gsub(">", "&gt;", value, fixed = TRUE)
  value
}

parameter_rows <- data.frame(
  group = c(
    rep("Trial size", 4),
    rep("Clinical thresholds", 3),
    rep("Credibility cutoffs", 4),
    rep("Calibration target", 5),
    rep("Scenario truth", 9),
    rep("Conditional effects", 2),
    rep("Dependence", 2),
    rep("Output", 3)
  ),
  parameter = c(
    "quick_mode",
    "dose_levels",
    "n_stages",
    "cohort_size",
    "phi_T",
    "phi_E",
    "phi_I",
    "c_T_start",
    "c_I_start",
    "c_E_start",
    "c_poc",
    "target_missing_range",
    "c_T_candidates",
    "c_I_candidates",
    "c_E_candidates",
    "n_sim_per_candidate",
    "high_tox_p_I",
    "high_tox_marginal_p_T",
    "high_tox_marginal_p_E",
    "low_immune_p_I",
    "low_immune_marginal_p_T",
    "low_immune_marginal_p_E",
    "low_eff_p_I",
    "low_eff_marginal_p_T",
    "low_eff_marginal_p_E",
    "toxicity_immune_effect",
    "efficacy_immune_effect",
    "rho0",
    "rho1",
    "append_history_log",
    "output_dir",
    "history_log_path"
  ),
  value = c(
    format_setting(quick_mode),
    format_setting(threshold_settings$dose_levels),
    format_setting(threshold_settings$n_stages),
    format_setting(threshold_settings$cohort_size),
    format_setting(threshold_settings$phi_T),
    format_setting(threshold_settings$phi_E),
    format_setting(threshold_settings$phi_I),
    format_setting(threshold_settings$c_T_start),
    format_setting(threshold_settings$c_I_start),
    format_setting(threshold_settings$c_E_start),
    format_setting(threshold_settings$c_poc),
    format_setting(threshold_settings$target_missing_range),
    format_setting(threshold_settings$c_T_candidates),
    format_setting(threshold_settings$c_I_candidates),
    format_setting(threshold_settings$c_E_candidates),
    format_setting(threshold_settings$n_sim_per_candidate),
    format_setting(threshold_settings$high_tox_p_I),
    format_setting(threshold_settings$high_tox_marginal_p_T),
    format_setting(threshold_settings$high_tox_marginal_p_E),
    format_setting(threshold_settings$low_immune_p_I),
    format_setting(threshold_settings$low_immune_marginal_p_T),
    format_setting(threshold_settings$low_immune_marginal_p_E),
    format_setting(threshold_settings$low_eff_p_I),
    format_setting(threshold_settings$low_eff_marginal_p_T),
    format_setting(threshold_settings$low_eff_marginal_p_E),
    format_setting(threshold_settings$toxicity_immune_effect),
    format_setting(threshold_settings$efficacy_immune_effect),
    format_setting(threshold_settings$rho0),
    format_setting(threshold_settings$rho1),
    format_setting(threshold_settings$append_history_log),
    format_path(threshold_settings$output_dir),
    format_path(threshold_settings$history_log_path)
  ),
  explanation = c(
    "Use TRUE for a fast smoke test. Use FALSE for a larger run once the setup looks right.",
    "Dose labels used by the adaptive design.",
    "How many times the trial analyzes accumulating data.",
    "Patients enrolled at each stage.",
    "Maximum acceptable marginal toxicity probability. Doses above this clinical limit should be treated as unsafe.",
    "Minimum acceptable marginal efficacy probability. Doses below this clinical limit should be treated as inactive.",
    "Minimum acceptable immune response probability. Doses below this clinical limit should be treated as biologically weak.",
    "Initial toxicity credibility cutoff. Calibration tests the candidate grid and may replace this value.",
    "Initial immune-response credibility cutoff. Calibration tests the candidate grid and may replace this value.",
    "Initial efficacy credibility cutoff. Calibration tests the candidate grid and may replace this value.",
    "Final PoC cutoff carried through this notebook for consistency; calibrate it in the PoC notebook.",
    "Target proportion of simulations where the final admissible set is empty in each unfavorable scenario.",
    "Grid searched for the toxicity credibility cutoff. Larger values require stronger safety evidence.",
    "Grid searched for the immune-response credibility cutoff. Larger values require stronger immune evidence.",
    "Grid searched for the efficacy credibility cutoff. Larger values require stronger efficacy evidence.",
    "Number of Monte Carlo trial simulations per candidate value.",
    "Immune response probability used while testing high toxicity.",
    "Target marginal toxicity probabilities used to make the c_T scenario unfavorable.",
    "Marginal efficacy probability kept acceptable while calibrating c_T.",
    "Immune response probabilities used to make the c_I scenario unfavorable.",
    "Marginal toxicity probability kept acceptable while calibrating c_I.",
    "Marginal efficacy probability kept acceptable while calibrating c_I.",
    "Immune response probability kept acceptable while calibrating c_E.",
    "Marginal toxicity probability kept acceptable while calibrating c_E.",
    "Target marginal efficacy probabilities used to make the c_E scenario unfavorable.",
    "How much P(T | I=1) differs from P(T | I=0), while preserving the target marginal P(T).",
    "How much P(E | I=1) differs from P(E | I=0), while preserving the target marginal P(E).",
    "Gumbel copula dependence parameter for toxicity and efficacy when I=0.",
    "Gumbel copula dependence parameter for toxicity and efficacy when I=1.",
    "Whether this run is appended to the readable Markdown calibration history.",
    "Folder where RDS and CSV calibration outputs are saved.",
    "Readable Markdown log that keeps a history of threshold calibration runs."
  ),
  stringsAsFactors = FALSE
)

cat(
  "<style>",
  "table:not(.parameter-guide-table) { display: block; max-width: 100%; overflow-x: auto; white-space: nowrap; }",
  "table:not(.parameter-guide-table) th, table:not(.parameter-guide-table) td { padding-left: 0.45rem; padding-right: 0.45rem; }",
  ".parameter-guide-table { width: 100%; border-collapse: collapse; table-layout: fixed; margin-bottom: 1.2rem; font-size: 0.9rem; line-height: 1.35; }",
  ".parameter-guide-table th, .parameter-guide-table td { border: 1px solid #d9d9d9; padding: 0.45rem 0.55rem; vertical-align: top; }",
  ".parameter-guide-table th { background: #f5f5f5; }",
  ".parameter-guide-table .param-col { width: 22%; font-family: monospace; overflow-wrap: anywhere; word-break: break-word; white-space: normal; }",
  ".parameter-guide-table .meaning-col { width: 53%; overflow-wrap: normal; word-break: normal; white-space: normal; }",
  ".parameter-guide-table .value-col { width: 25%; overflow-wrap: anywhere; word-break: break-word; white-space: normal; }",
  "</style>\n",
  sep = ""
)

cat("<p><strong>User settings and parameter meanings.</strong> Values are summarized when a grid is long; the full grid is still defined in the User Settings chunk above.</p>\n")

for (group_name in unique(parameter_rows$group)) {
  group_rows <- parameter_rows[parameter_rows$group == group_name, , drop = FALSE]
  cat("<h3>", escape_html(group_name), "</h3>\n", sep = "")
  cat("<table class=\"parameter-guide-table\">\n")
  cat("<thead><tr><th class=\"param-col\">Parameter</th><th class=\"meaning-col\">Meaning</th><th class=\"value-col\">Current value</th></tr></thead>\n")
  cat("<tbody>\n")
  for (row_index in seq_len(nrow(group_rows))) {
    row <- group_rows[row_index, , drop = FALSE]
    cat(
      "<tr><td class=\"param-col\">",
      escape_html(row$parameter),
      "</td><td class=\"meaning-col\">",
      escape_html(row$explanation),
      "</td><td class=\"value-col\">",
      escape_html(row$value),
      "</td></tr>\n",
      sep = ""
    )
  }
  cat("</tbody></table>\n")
}
```

## 4. Scenario Check

The notebook builds three unfavorable scenarios from the user settings. Conditional toxicity and efficacy matrices are calculated from the target marginal probabilities and immune-response probabilities.

```{r scenario_check}
threshold_scenarios <- list(
  c_T = create_threshold_scenario("toxicity", threshold_settings),
  c_I = create_threshold_scenario("immune", threshold_settings),
  c_E = create_threshold_scenario("efficacy", threshold_settings)
)

scenario_table <- bind_rows(lapply(names(threshold_scenarios), function(name) {
  scenario <- threshold_scenarios[[name]]
  data.frame(
    calibration = name,
    scenario = scenario$description,
    p_I = paste(round(scenario$p_YI, 3), collapse = ", "),
    marginal_p_T = paste(round(scenario$marginal_p_T, 3), collapse = ", "),
    marginal_p_E = paste(round(scenario$marginal_p_E, 3), collapse = ", "),
    stringsAsFactors = FALSE
  )
}))

kable(scenario_table, caption = "Endpoint-specific calibration scenarios")
```

## 5. Run Calibration

```{r run_calibration}
calibration_results <- calibrate_separate_thresholds(threshold_settings)
summary_table <- threshold_calibration_summary_table(calibration_results)

display_summary <- summary_table %>%
  mutate(
    selected_value = round(selected_value, 3),
    achieved_missing_rate = paste0(round(achieved_missing_rate * 100, 1), "%"),
    target = paste0(round(target_low * 100), "%-", round(target_high * 100), "%")
  ) %>%
  select(parameter, endpoint, selected_value, achieved_missing_rate, target, status, n_simulations)

kable(display_summary, caption = "Recommended threshold values")
```

```{r candidate_tables}
#| results: asis

for (name in names(calibration_results$calibrations)) {
  result <- calibration_results$calibrations[[name]]
  candidate_table <- result$results %>%
    transmute(
      parameter = param_name,
      candidate = round(param_value, 3),
      final_missing_rate = paste0(round(final_admissible_missing_rate * 100, 1), "%"),
      target_endpoint_missing_rate = paste0(round(target_endpoint_missing_rate * 100, 1), "%"),
      early_stop_rate = paste0(round(early_stop_rate * 100, 1), "%"),
      mean_admissible_count = round(mean_admissible_count, 2)
    )
  cat("\n\n### Candidate Results for ", name, "\n\n", sep = "")
  print(kable(candidate_table))
}
```

## 6. Plots

```{r threshold_plots, fig.width=8, fig.height=5}
plot_threshold_calibration(calibration_results$calibrations$c_T)
plot_threshold_calibration(calibration_results$calibrations$c_I)
plot_threshold_calibration(calibration_results$calibrations$c_E)
```

## 7. Save Outputs

```{r save_outputs}
dir.create(threshold_settings$output_dir, showWarnings = FALSE, recursive = TRUE)

rds_path <- file.path(threshold_settings$output_dir, "threshold_calibration_results.rds")
csv_path <- file.path(threshold_settings$output_dir, "threshold_calibration_summary.csv")

saveRDS(calibration_results, rds_path)
write.csv(summary_table, csv_path, row.names = FALSE)

history_path <- NA_character_
if (isTRUE(threshold_settings$append_history_log)) {
  history_path <- append_threshold_calibration_log(
    calibration_results,
    file_path = threshold_settings$history_log_path,
    run_label = threshold_settings$run_label
  )
}

cat("Saved RDS:", rds_path, "\n")
cat("Saved summary CSV:", csv_path, "\n")
if (!is.na(history_path)) {
  cat("Appended history log:", history_path, "\n")
}
```

## 8. Use These Values Next

```{r final_values}
recommended <- calibration_results$recommended_thresholds

cat("Recommended threshold settings:\n")
cat("  c_T =", recommended$c_T, "\n")
cat("  c_I =", recommended$c_I, "\n")
cat("  c_E =", recommended$c_E, "\n")
cat("\nNext step: run the PoC calibration notebook. By default it reads the saved threshold_calibration_results.rds file and uses these values before calibrating c_poc.\n")
```
