---
title: "Threshold Calibration Notebook"
subtitle: "Separate calibration of c_T, c_I, and c_E before PoC calibration"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show Code"
    embed-resources: true
    self-contained: true
execute:
  message: false
  warning: false
  echo: false
editor: visual
---

## 1. Purpose

This notebook calibrates the three admissibility credibility cutoffs separately:

1. `c_T`: evidence required to keep a dose when toxicity is judged acceptable.
2. `c_I`: evidence required to keep a dose when immune response is judged acceptable.
3. `c_E`: evidence required to keep a dose when efficacy is judged acceptable.

The calibration target is the final admissible set missing rate: the proportion of simulated trials where the final admissible set is empty under an endpoint-specific unfavorable scenario. The default target is 80%-90%.

After this notebook selects `c_T`, `c_I`, and `c_E`, use those values as fixed inputs in the PoC calibration notebook.

## User Settings

Edit this chunk only for routine use.

```{r user_settings}
#| echo: true

quick_mode <- TRUE

threshold_settings <- list(
  quick_mode = quick_mode,

  dose_levels = c(1, 2, 3, 4, 5),
  n_stages = 5,
  cohort_size = 15,

  phi_T = 0.30,
  phi_E = 0.25,
  phi_I = 0.20,

  c_T_start = 0.55,
  c_E_start = 0.50,
  c_I_start = 0.70,
  c_poc = 0.995,
  delta_poc = 0.8,
  rho0 = 1.5,
  rho1 = 2.0,

  target_missing_range = c(0.80, 0.90),
  c_T_candidates = if (quick_mode) c(0.45, 0.55, 0.65) else seq(0.35, 0.75, by = 0.05),
  c_I_candidates = if (quick_mode) c(0.50, 0.70, 0.90) else seq(0.45, 0.95, by = 0.05),
  c_E_candidates = if (quick_mode) c(0.35, 0.50, 0.65) else seq(0.30, 0.80, by = 0.05),
  n_sim_per_candidate = if (quick_mode) 5 else 500,
  calibration_seed = 30000,

  high_tox_p_I = 0.30,
  high_tox_marginal_p_T = c(0.35, 0.60),
  high_tox_marginal_p_E = 0.40,

  low_immune_p_I = c(0.10, 0.15),
  low_immune_marginal_p_T = 0.15,
  low_immune_marginal_p_E = 0.40,

  low_eff_p_I = 0.30,
  low_eff_marginal_p_T = 0.15,
  low_eff_marginal_p_E = c(0.10, 0.20),

  toxicity_immune_effect = 0,
  efficacy_immune_effect = 0,

  output_dir = "results/threshold_calibration",
  append_history_log = TRUE,
  history_log_path = file.path("results", "threshold_calibration", "threshold_calibration_history.md"),
  run_label = if (quick_mode) "quick smoke test" else "production calibration"
)
```

## 2. Setup

```{r setup}
#| message: false
#| warning: false
#| echo: false

library(dplyr)
library(ggplot2)
library(knitr)

project_root_candidates <- c(".", "..", "../..")
project_root_matches <- project_root_candidates[
  file.exists(file.path(project_root_candidates, "DoseFinding.Rproj"))
]
if (length(project_root_matches) == 0) {
  stop("Could not find project root containing DoseFinding.Rproj.")
}
project_root_abs <- normalizePath(project_root_matches[[1]], winslash = "/", mustWork = TRUE)
setwd(project_root_abs)

source("src/optimization/threshold_calibration.R")

threshold_settings$output_dir <- file.path(project_root_abs, threshold_settings$output_dir)
threshold_settings$history_log_path <- file.path(project_root_abs, threshold_settings$history_log_path)
```

## 3. Scenario Check

The notebook builds three unfavorable scenarios from the user settings. Conditional toxicity and efficacy matrices are calculated from the target marginal probabilities and immune-response probabilities.

```{r scenario_check}
threshold_scenarios <- list(
  c_T = create_threshold_scenario("toxicity", threshold_settings),
  c_I = create_threshold_scenario("immune", threshold_settings),
  c_E = create_threshold_scenario("efficacy", threshold_settings)
)

scenario_table <- bind_rows(lapply(names(threshold_scenarios), function(name) {
  scenario <- threshold_scenarios[[name]]
  data.frame(
    calibration = name,
    scenario = scenario$description,
    p_I = paste(round(scenario$p_YI, 3), collapse = ", "),
    marginal_p_T = paste(round(scenario$marginal_p_T, 3), collapse = ", "),
    marginal_p_E = paste(round(scenario$marginal_p_E, 3), collapse = ", "),
    stringsAsFactors = FALSE
  )
}))

kable(scenario_table, caption = "Endpoint-specific calibration scenarios")
```

## 4. Run Calibration

```{r run_calibration}
calibration_results <- calibrate_separate_thresholds(threshold_settings)
summary_table <- threshold_calibration_summary_table(calibration_results)

display_summary <- summary_table %>%
  mutate(
    selected_value = round(selected_value, 3),
    achieved_missing_rate = paste0(round(achieved_missing_rate * 100, 1), "%"),
    target = paste0(round(target_low * 100), "%-", round(target_high * 100), "%")
  ) %>%
  select(parameter, endpoint, selected_value, achieved_missing_rate, target, status, n_simulations)

kable(display_summary, caption = "Recommended threshold values")
```

```{r candidate_tables}
#| results: asis

for (name in names(calibration_results$calibrations)) {
  result <- calibration_results$calibrations[[name]]
  candidate_table <- result$results %>%
    transmute(
      parameter = param_name,
      candidate = round(param_value, 3),
      final_missing_rate = paste0(round(final_admissible_missing_rate * 100, 1), "%"),
      target_endpoint_missing_rate = paste0(round(target_endpoint_missing_rate * 100, 1), "%"),
      early_stop_rate = paste0(round(early_stop_rate * 100, 1), "%"),
      mean_admissible_count = round(mean_admissible_count, 2)
    )
  cat("\n\n### Candidate Results for ", name, "\n\n", sep = "")
  print(kable(candidate_table))
}
```

## 5. Plots

```{r threshold_plots, fig.width=8, fig.height=5}
plot_threshold_calibration(calibration_results$calibrations$c_T)
plot_threshold_calibration(calibration_results$calibrations$c_I)
plot_threshold_calibration(calibration_results$calibrations$c_E)
```

## 6. Save Outputs

```{r save_outputs}
dir.create(threshold_settings$output_dir, showWarnings = FALSE, recursive = TRUE)

rds_path <- file.path(threshold_settings$output_dir, "threshold_calibration_results.rds")
csv_path <- file.path(threshold_settings$output_dir, "threshold_calibration_summary.csv")

saveRDS(calibration_results, rds_path)
write.csv(summary_table, csv_path, row.names = FALSE)

history_path <- NA_character_
if (isTRUE(threshold_settings$append_history_log)) {
  history_path <- append_threshold_calibration_log(
    calibration_results,
    file_path = threshold_settings$history_log_path,
    run_label = threshold_settings$run_label
  )
}

cat("Saved RDS:", rds_path, "\n")
cat("Saved summary CSV:", csv_path, "\n")
if (!is.na(history_path)) {
  cat("Appended history log:", history_path, "\n")
}
```

## 7. Use These Values Next

```{r final_values}
recommended <- calibration_results$recommended_thresholds

cat("Recommended threshold settings:\n")
cat("  c_T =", recommended$c_T, "\n")
cat("  c_I =", recommended$c_I, "\n")
cat("  c_E =", recommended$c_E, "\n")
cat("\nNext step: copy these values into the PoC calibration notebook User Settings before calibrating c_poc.\n")
```
